---
title: "Messer-Sealey-White-ADA-2020-vignette-assignment"
author: "BA White, BA Sealey, E Messer"
date: "5/13/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Objectives
The objective of this module is to provide an overview to utilizing network analysis to visualize and analyze interaction, connected, or similarity data. Specifically, we present 1. Tools to modify a data matrix to generate network diagrams, 2. Applying various aesthetics to visNetwork diagrams, and 3. Generate Wilcoxon Test statistics make comparisons across groups.

# Preliminaries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # to enable tidying of data
library(reshape2) # to use with melt function to convert matrix to objects igraph can plot
library(igraph) # to make an igraph object ie. network visualization of matrix values
library(visNetwork) # to make your igraph object prettier using visNetwork
library(bbe2020) # adds functionality to melt by converting a matrix generated in other softwares or elsewhere and converting to a form easily used by igraph
```

# Introduce dataset 
A group of snakes were exposed to the presence of two "simulated" predators  (ie. two treatments), one with a simulated potential aerial "attack" and another with a simulated tactile "attack". A matrix was generated using Pearson correlation coefficient to calculate the behavior similarity indices between all snake's antipredator head movement behavior.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# read in data matrix for aerial treatment
aerial <- read_csv("blacktreatments_heads2020f.csv", col_names=TRUE)
# store the first column, which should be "names" 
names <- as.character(aerial$names)
# read in data matrix for tactile treatment 
tactile <- read_csv("tactiletreatments_heads2020f.csv", col_names=TRUE)
# store the first column, which should be "names" 
names <- as.character(tactile$names)

## Please Note: you must ensure that your dataframe has a first column named "names" that refers to the labels for the unique row and column data points 
```

# Specific question
How similar did individual vipers move their heads in response to an aerial "attack"?
How similar did individual vipers move their heads in response to a tactile "attack"?
How does viper centrality of similarity compare between the aerial and tactile treatments? 

# Background
Networks are made up of two essential components, 'nodes' (represented by symbols) and ‘edges’ (represented by lines between the nodes). These lines typically indicate an interaction between the nodes. Usually, in the context of social networks, the nodes are individuals and the edges are some measured interaction (Croft, James & Krause, 2008). Social network visualizations are called sociograms and illustrate a graphical representation of the data (presented as nodes and edges). Network visualizations can be used to visualize many forms of interaction profile similarity or adjacency matrices in biological systems (Bass et al. 2013).   

# Tidying Data for analysis  
```{r}
# melting data matrix for aerial treatment
colnames(aerial) <-  c("names", 1:length(aerial))
cleanaerial <- select(aerial, -names)
head(cleanaerial)
cleanaerial <- as.matrix(cleanaerial)
cleanaerial[lower.tri(cleanaerial, diag=TRUE)] <- NA
meltaerial <- setNames(melt(cleanaerial), c('ind1', 'ind2', 'values'))
head(meltaerial)
meltaerial <- filter(meltaerial, !is.na(values))
summary(meltaerial)

meltmeaerial <- meltme(aerial)
meltmeaerial
# melting data matirx for tactile treatment 
d <- read_csv("tactiletreatments_heads2020f.csv", col_names=TRUE)
colnames(d) <-  c("names", 1:length(d))
d <- select(d, -names)
head(d)
d <- as.matrix(d)
d[lower.tri(d, diag=TRUE)] <- NA
r1 <- setNames(melt(d), c('ind1', 'ind2', 'values'))
head(r1)
r1 <- filter(r1, !is.na(values))
summary(r1)
```

# Data Visualization
Here we specifically visualize an individual snake's antipredator behavior divided by all individual's antipredator behavior towards the same type of predator represented as a ratio (or similarity index) in a sociogram. Each sociogram visualizes the similarity of antipredator behavior for either predator among the snakes. Edges represent the similarity index. Centrality is represented by node size.

```{r}
# network plot of aerial treatment
aerialplot <- graph_from_data_frame(meltaerial, directed = FALSE, vertices = NULL)
plot(aerialplot, vertex.label = NA, layout = layout_on_sphere, main="Aerial treatment") # using a force-directed layout (positions the nodes so that the edges have similar lengthes and there are as few crossing edges as possible)

# network plot of tactile treatment
p1 <- graph_from_data_frame(r1, directed = FALSE, vertices = NULL)
plot(p1, vertex.label = NA, layout = layout_with_fr, main="Tactile treatment")  # using a force-directed layout (positions the nodes so that the edges have similar lengthes and there are as few crossing edges as possible)

library(visNetwork)
# converting igraph to visnetwork for aerial treatment
visIgraph(aerialplot, randomSeed = TRUE) %>% 
  visOptions(manipulation = TRUE) 

# converting igraph to visnetwork for tactile treatment
visIgraph(p1, randomSeed = TRUE) %>% 
  visOptions(manipulation = TRUE) 

# visnetwork for aerial treatment with new layout
data <- toVisNetworkData(aerialplot)

## copy column "weight" to new column "value" in list "edges"
data$edges$width <- 20*data$edges$value # line width
data$edges$color <- "gray" # line color
data$nodes$label <- names
visNetwork(nodes = data$nodes, edges = data$edges, height = "500px") %>%
visIgraphLayout(layout = "layout_with_sugiyama")

visNetwork(nodes = data$nodes, edges = data$edges, height = "500px") %>%
visIgraphLayout(layout = "layout_with_sugiyama") %>% 
  visOptions(highlightNearest = list(enabled =TRUE, degree = 20))

# URL FOR NODE WEIGHT POT https://cran.r-project.org/web/packages/visNetwork/vignettes/Introduction-to-visNetwork.html

```

# Compare centrality of similarity index between an aerial and tactile treatments. 
```{r}

## EM updating this seperately... 

# eigenvector centrality for snakes in the aerial treatment
cent_aerial <- (eigen_centrality(aerialplot))
# eigenvector centrality for snakes in the tactile treatment
cent_tact <- (eigen_centrality(p1))  
```

# Statistical Approach
To determine if there is any significant difference between the predator responses of our snakes towards the two artificially simulated predators (aerial and tactile) we compared the snakes scores of centrality (connectedness). Centrality is a network statistic which is a measure of being well-connected.
```{r}
# First we need to combine our centrality measures for both the aerial and tactical conditiosn into one dataframe, so 3 columns: id, treatment (aerial or tactile), centrality score and add in the name of the dataframe in the viper_wilcox analysis below and this should (hopefully) work...

#insert new dataframe here 


# Wilcoxon Test to compare the centrality scores for the vipers in the two different treatments 
viper_wilcox <- # dataname  %>& 
  wilcox_test(centrality ~ treatment, paired = TRUE, p.adjust.method = "none")
viper_wilcox

```


# Another Example Using Association to Demonstarte Aesthetic Changes 
In this new example, we showcase a selection of different aesthetic changes you can utilize in your visualization of network data based on monkey association patterns. 
```{r}
# Monkeys also work 
Monkeys <- read_csv("EM_monkey.csv", col_names=TRUE)
colnames(Monkeys) <-  c("name", 1:length(Monkeys))
Monkeys <- select(Monkeys, -name)
Monkeys <- as.matrix(Monkeys)
Monkeys[lower.tri(Monkeys, diag=TRUE)] <- NA 
Monkeys <- setNames(melt(Monkeys), c('Source', 'Target', 'Weight'))
Monkeys <- filter(Monkeys, !is.na(Weight))
summary(Monkeys)

melt_matrix_4_visNetwork <- function(z){
  colnames(x) <- c("names", 1:length(x))
  x <- select(x, -names) %>%
  as.matrix()
  x[lower.tri(x, diag=TRUE)] <- NA
  x <- melt(x)
  x <- setNames(x, c('ind1', 'ind2', 'values'))
  x <- filter(x, !is.na(values))
  return(x)
}

new <- melt_matrix_4_visNetwork(Monkeys)  
head(new)


# This is how you can replace numbers with names 
Sname <- c(Monkeys$name)
Sno <- as.numeric(1:nrow(Monkeys))
Source <- bind_cols(name = Sname, number = Sno)
Source

final <- inner_join(new, Source, by = c("ind1" = "number"))
final <- inner_join(final, Source, by = c("ind2" = "number"))

final <- select(final, name.x, name.y, values)
final

# Visualization of Monkey interactions 
Monkeysp <- graph_from_data_frame(final, directed = FALSE, vertices = NULL)
plot(aerialplot, vertex.label = NA, layout = layout_with_fr, main="Aerial treatment") 

datam <- toVisNetworkData(Monkeysp)
visNetwork(nodes = datam$nodes, edges = datam$edges, height = "500px") %>%
visIgraphLayout(layout = "layout_with_sugiyama")

# New monkey graph: 
#igraph
monk = graph.data.frame(final[,c('name.x','name.y')], directed = FALSE, vertices = NULL)
E(monk)$weight = final$values
head(monk)
plot(monk, vertex.label = NA, layout = layout_with_fr, main="Monkeys") 

#then in visNetwrk 
datam <- toVisNetworkData(monk)
visNetwork(nodes = datam$nodes, edges = datam$edges, height = "500px") %>%
  visIgraphLayout(layout = "layout_with_sugiyama")

# working out centrality 
centrality <- evcent(monk)$vector
head(centrality) 

````

# Inferring Results 
Individual snakes typically behaved significantly different between treatments as expected. 

# Concept Overview 
Networks are made up of two essential components, 'nodes' (represented by symbols) and ‘edges’ (represented by lines between the nodes). These lines typically indicate an interaction between the nodes
Using the melt function bbe2020() to convert matrix data into a dataframe for network visualization & analysis
Using {igraph}: graph_from_data_frame() and {visNetwork}: toVisNetworkData() to generate sociograms 
Aesthetic changes in visNetwork sociograms <add in brief code example here>
Centrality statistics in igraph evcent()
The wilcox_test() is a non-parametric statistical test which compares two paired groups 
