---
title: "Messer-Sealey-White-ADA-2020-vignette-assignment"
author: "BA White, BA Sealey, E Messer"
date: "5/13/2020"
output: html_document
---
## Files to Turn In: 
## 1) RMD file(1) and thus html(2) for vignette module
## 2) vignette R package (3) using 1 or more dataset and NOW 2 functions of choice 

## Vignette Assignemt Objectives:
## 1a) walk users through demonstrations of relevant analyses in module style 
## 2a) set of functions and associated function documentation appropriate to the topic you have chosen
## 2b) one or more relevant datasets (either data you have simulated or, preferably, real data), that you use in a module vignette about your chosen topic
## 2c) A vignette that walks users through relevant analyses and appropriate METADATA for your package (e.g., information on dependencies, etc.)

# Objectives
The objective of this module is to provide an overview to utilising social network analsyis in visualizing and analysing interaction data. Specifically we present 1. Tools to create and modify network diagrams, 2. methods to generate social network statsistics and make comparisons across groups... - come back here when we work out what to include...  

# Preliminaries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # to enable tidying of data
library(reshape2) # to use with melt function to convert matrix to objects igraph can plot
library(igraph) # to make an igraph object ie. network visualization of matrix values. 

# more libraries etc
```

# Introduce NEW dataset 
Snakes were exposed to the presence of two "simulated" predators, were an individual was exposed to two treatments, one with a simulated avian predator and another with a tactile predator touch. Dataset includes a matrix of behavior similarity index between all snake's antipredator behavior.    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# read in data matrix for aerial treatment
aerial <- read_csv("blacktreatments_heads2020f.csv", col_names=TRUE)

# read in data matrix for tactile treatment 
tactile <- read_csv("tactiletreatments_heads2020f.csv", col_names=TRUE)

## Please Note: you must ensure that your dataframe has a first column named "names" that refers to unique row and column data points 
```

# Specific question
## How similar did individual vipers move their heads in response to an aerial "attack"?
## How similar did individual vipers move their heads in response to a tactile "attack"?
## How does viper mean centrality of similarity compare between the aerial and tactile treatments? 

# Background
Networks are made up of two essential components, 'nodes' (represented by symbols) and ‘edges’ (represented by lines between the nodes). These lines typically indicate an interaction between the nodes. Usually, in the context of social networks, the nodes are individuals and the edges are some measured interaction (Croft, James & Krause, 2008). Social network visualizations are called sociograms and illustrate a graphical representation of the data presented as nodes and edges. But network visualizations can be used to visulize many forms of interaction profile similarity or adjacency matrices in biologicial systems (Bass et al. 2013). Here we specifically visualize an individual snake's antipredator behavior divided by all individual's antipredator behavior represented as a ratio in a network of similarity. This visualizes the similarity of antipredator behavior among the snakes. We then determine if the centrality of the similarity ratio between two treatments, aerial and tactile, differs.  

# Tidying Data for analysis  
```{r}
# melting data matrix for aerial treatment
b0 <- read_csv("blacktreatments_heads2020f.csv", col_names=TRUE)
colnames(aerial) <-  c("names", 1:length(b0))
cleanaerial <- select(aerial, -names)
head(cleanaerial)
cleanaerial <- as.matrix(cleanaerial)
cleanaerial[lower.tri(cleanaerial, diag=TRUE)] <- NA
meltaerial <- setNames(melt(cleanaerial), c('ind1', 'ind2', 'values'))
head(meltaerial)
meltaerial <- filter(meltaerial, !is.na(values))
summary(meltaerial)

# melting data matirx for tactile treatment 
d <- read_csv("tactiletreatments_heads2020f.csv", col_names=TRUE)
colnames(d) <-  c("names", 1:length(d))
d <- select(d, -names)
head(d)
d <- as.matrix(d)
d[lower.tri(d, diag=TRUE)] <- NA
r1 <- setNames(melt(d), c('ind1', 'ind2', 'values'))
head(r)
r1 <- filter(r1, !is.na(values))
summary(r1)
```


# Data Visualization
```{r}
# network plot of aerial treatment
aerialplot <- graph_from_data_frame(meltaerial, directed = FALSE, vertices = NULL)
plot(aerialplot, vertex.label = NA)

# network plot of tactile treatment
p1 <- graph_from_data_frame(r1, directed = FALSE, vertices = NULL)
plot(p1, vertex.label = NA)


library(visNetwork)
# converting igraph to visnetwork for aerial treatment
visIgraph(aerialplot, randomSeed = TRUE) %>% 
  visOptions(manipulation = TRUE)

# converting igraph to visnetwork for tactile treatment
visIgraph(p1, randomSeed = TRUE) %>% 
  visOptions(manipulation = TRUE)


```

# Our prediction is that the visualization will cluster antipredator behavior in the center - lines represent the similarity by thickness. Centrality is represented by node size. 

# Statistical Approach
# Compare similarity index between an aerial and tactile treatments. 

# Inferring Results 
# Individual snakes typically behaved similarly in both treatments as expected. 

# Concept Overview 

